---
title: "Brandt C3.4 HW"
author: "Kate Brandt"
date: "January 18, 2019"
output:
  word_document: default
  html_document: default
---
This includes my code script (I completed the assignment in R) and my answers

**Setting up**
``` {r, warning=FALSE, message=FALSE, results = 'hide'}
# set workspace for SOCI709
setwd("~/R/SOCI709/analysis")

# import necessary packages to import .dta file
library(foreign)
library(tidyverse)

# import data and create variables
attend <-read.dta("data/attend.dta")
atndrte <- attend$atndrte
priGPA <- attend$priGPA
ACT <- attend$ACT

```


**(i) Min, max, mean for selected variables** 

  atndrte

  * Min: 6.25
  * Max: 100
  * Mean: 81.70956
  
  priGPA
  
  * Min: 0.857
  * Max: 3.93
  * Mean: 2.58775
  
  ACT
  
  * Min: 133
  * Max: 32
  * Mean: 22.51029
  
``` {r, warning=FALSE, message=FALSE, results = 'hide'}
# Code for (i) to find Min, Max, Mean
i_funs <- c("min", "max", "mean")
attend %>% select(atndrte, priGPA, ACT) %>% summarise_all(funs_(i_funs))
```


**(ii) Estimate model**

Estimated Model:
  
  atndrte = 75.7 + 17.261(priGPA) - 1.717(ACT) + u
  
  Interpretation: 
  
  
```{r, warning=FALSE, message=FALSE, results = 'hide'}
# Code for (ii) to estimate model
## model: atndrte = B0 + b1(priGPA) + b2(ACT) + u

## set variables 
head(attend)
y<- (atndrte)
x1<- (priGPA)
x2<- (ACT)

## Create equation and summarize
attend <-
  cbind.data.frame(attend,y,x1,x2)
head(attend)
mod_ii <- lm(y~x1+x2, data=attend)
summary(mod_ii)

```

**(iii)**



**(iv) Using estimated model to predict attendance rate**
  
  Predicted attendance rate: 104.3627%
  
  Interpretation: This is an unrealistic result- nobody can have over a 100% attendance rate. 

```{r, warning=FALSE, message=FALSE, results = 'hide'}
# Code for (iv) to calculate priGPA = 3.65 and ACT = 20
atndrte_iv <- 75.7 + 17.261*3.65 - 1.717*20
atndrte_iv

```


**(v) Calculating differences between predicted values**

  Predicted attendance rate for student A: 93.1521%
  
  Predicted attendance rate for student B: 67.3061%
  
  Difference between rates: 25.846%
  
  
  Mathematically, this difference makes sense because of the negative coefficient associated with ACT score and higher emphasis on GPA value in this model.
  
```{r, warning=FALSE, message=FALSE, results = 'hide'}
# Code for (v) to predict values and take the difference
#predict and print values
studentA <- 75.7 + 17.261*3.1 - 1.717*21
studentA
studentB <- 75.7 + 17.261*2.1 - 1.717*26
studentB
#calculate difference
diffAB <- studentA - studentB
diffAB
```

